---
output:
    html_document:
        keep_md: yes
---
---
layout: post
author: "csiu"
date: 2016-06-05
title: Perfectionism
categories: update
tags: data-analysis
excerpt:
header-img-http: https://pixabay.com/static/uploads/photo/2015/02/13/14/38/pretty-woman-635258_960_720.jpg
header-img-source: jil111@pixabay
---
```{r include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      #results='hide',
                      #fig.path='/Users/csiu/project/blog/myblog/img/figure/YYYY-MM-DD',
                      fig.retina = 2
		      )
options(knitr.table.format = 'markdown')
```

```{r load-packages, message=FALSE, warning=FALSE}
## Reformatting data
library(tidyr)
library(dplyr)

## Visualization
library(ggplot2)
library(cowplot)

theme_set(theme_bw())
```
<style>
blockquote p {
    margin: 0;
    padding: 0;
    font-family: 'Open Sans', 'Helvetica Neue', Helvetica, Arial, sans-serif;
    font-style: normal;
    color: silver;
}
blockquote li {
    font-family: "Courier New", Courier, monospace;
    font-style: normal;
    color: silver;
}
</style>

Three weeks ago, I -- along with 34 other people -- attended an “Overcoming perfectionism: Bring more ease & success into your life” workshop.

> **2 futile formulas:**
>
> - Self worth = Ability = Performance
> - Ordinary = Not outstanding = Not loved

Seeing as there were more females than males, I wondered (assuming a null hypothesis of equal turnout of males and females) whether there was significantly more females than males present at this event.

> **3 types of perfectionism:**
>
> - Self-orientated
> - Other-orientated
> - Socially-orientated

Applying the [exact binomial test of goodness-of-fit](http://www.biostathandbook.com/exactgof.html) to test for equal attendance of females and males, we get a p-value (representing the probability of getting a result like the observed result if the null hypothesis were true) of 0.04.

```{r}
binom.test(24, 35, 0.5)
```

> **3 types of thinking traps:**
>
> - All or nothing thinking
> - Should/Must thinking
> - Minimization/Magnification


Seeing that the p-value of 0.04 is less than the typical significant level of 0.05, we reject the null hypothesis and conclude that there were significantly more females than males present at this event. 
