---
output:
    html_document:
        keep_md: yes
---
---
layout: post
author: "csiu"
date: 2016-05-22
title: "Rating Noodler's Ink"
categories: update
tags: data-analysis
excerpt: "I find myself wanting a new ink for my fountain pen, but the problem I have is that I can't decide on which one."
header-img-http: https://pixabay.com/static/uploads/photo/2014/09/15/22/23/fountain-pen-447576_960_720.jpg
header-img-source: Andrys@pixabay
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      #results='hide',
                      fig.retina = 2,
                      fig.path='/Users/csiu/project/blog/myblog/img/figure/2016-05-22/'
                      )
options(knitr.table.format = 'markdown')
```

```{r load-packages, message=FALSE, warning=FALSE}
## Reformatting data
library(tidyr)
library(dplyr)
library(readr)

## Visualization
library(ggplot2)
library(viridis)
library(cowplot)

theme_set(theme_bw())
```

> A human being is a deciding being.<br>
> -- Viktor E. Frank

I find myself wanting a new ink for my fountain pen, but the problem I have is that I can't decide on which one.

```{r load-data}
inks <- read_csv("noodersinkproperties.csv")
ranks <- read_csv("goulet-ranks.csv")
```

Looking at a [2012 comparison of ink prices per volume](http://www.marcuslink.com/pens/ink/thebrands.html), Noodler's ink would be the most economical choice. In this line, there are [`r nrow(inks)` different inks with various ink properties](http://noodlersink.com/noodlers-ink-properties/) such as:

- **Bulletproof**: whether the ink is resistant to UV, industrial solvents, or bleaching agents
- **Eternal**: whether the ink is fade resistant -- good for archival purposes
- **Fluorescent**: whether the ink glows under UV light
- **Forgery Resistant**: whether the ink is impervious to lasers, alcohols, or solvents -- ideal for security documents
- **Freeze Resistant**: whether the ink resists freezing at extremely low temperatures
- **Lubricated**: whether the ink is lubricated (for use in piston mechanism)
- **Water Resistant**: whether the ink is waterproof

## Different inks have different properties

```{r plot-ink-property, warning=FALSE}
inks %>% 
  gather(property, value, -Ink) %>% 
  ggplot(aes(x = value)) +
    geom_bar() +
    facet_grid(~property, scales = "free_x", space = "free_x") +
    xlab("") +
    ggtitle("Properties of Noodler's ink") +
    theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))
```

```{r number-of-ink-property-combos}
nprop <- apply(inks, 2, function(x){x %>% na.omit() %>% unique() %>% length()})
nprop <- nprop[inks %>% select(-Ink) %>% colnames()]
```

In the plot, we can see the different values a property can take; and if we consider all possible combinations, there will be `r nprop[order(names(nprop))] %>% paste(collapse="x")` = `r prod(nprop)` possibilities. Fortunately, I am able to narrow down the combination of properties to `r 3*3*3` choices, for I want an ink that:

1. is (or at least is partially) bulletproof,
2. will not [fade](http://hudsonvalleysketches.blogspot.ca/2012/12/lightfastness-test-results-on-noodlers.html), and
3. is (or at least is partially) waterproof

```{r shortlist-inks}
inks_subset <- inks %>% 
  filter(
    Eternal != "No",
    Bulletproof != "No",
    Water_Resistant != "No"
  ) 
```

## Shortlist of 43 inks

After filtering Noodler's list of `r nrow(inks)` inks, I am left with `r nrow(inks_subset)` matching inks.

<font color="silver"><em>`r gsub("^\\#", '&#35;', inks_subset$Ink) %>% paste(collapse=", ")`</em></font>

This is still a lot. I now need a way to prioritize this list and one way to do so is to consider [The Goulet Pen](http://www.gouletpens.com/bottled-ink/c/14/?sortBy=productName+asc&facetValueFilter=Tenant~Brand%3Anoodlers) [ratings \(taken on May 22, 2016\)](https://gist.github.com/csiu/7d53c19edc8ff10798aa773ac3508974).

```{r tidy-ranks}
ranks <- ranks %>% 
  separate(name, into=c("name", "volume"), sep=" \\(") %>% 
  mutate(
    has_free_pen = ifelse(!grepl("\\)$", volume), T, F),
    volume = sub("oz.*\\).*", "", volume) %>% as.numeric(),
    name = sub("^Noodler's ", "", name)
  )
```

## Ink ratings from The Goulet Pen

```{r dups}
dups <- ranks$name[ranks$name %>% duplicated()]
dups <- ranks %>% filter(name %in% dups)
```

On the website, Goulet lists `r nrow(ranks)` bottles of Noodler's inks, `r dups$name %>% unique() %>% length()` of which are duplicated in 3oz and 4.5oz options.

```{r plot-dups, fig.height=4}
## Plot duplicate inks
dups %>% 
  mutate(volume = sprintf("%s oz", volume)) %>% 
  ggplot(aes(x=reorder(name, rating), 
             y=rating, 
             color=volume, 
             group=volume)) +
  geom_point(alpha=.7) + 
  geom_line(alpha=.7) +
  xlab("ink") +
  ggtitle("Ratings of inks available in both the 3oz & 4.5oz options") +
  theme(axis.text.x = element_blank())
```

```{r binom}
## Binomial test
run_binomtest <- function(x){
  tbl <- table(x)
  binom.test(x = tbl[["TRUE"]],
             n = sum(tbl),
             p = 0.5,
             alternative = "two.sided")
}

result <- dups %>% 
  select(name, volume, rating) %>% 
  spread(volume, rating) %>% 
  mutate(
    is_4_higher_rating = `4.5` > `3`,
    is_4_higherEqual_rating= `4.5` >= `3`
    ) %>% 
  {.$is_4_higher_rating} %>% 
  run_binomtest()
result$data.name <- "A success is counted if the rating is higher in the 4.5oz group"
```

When tested for whether the ratings were higher in 4.5oz option vs 3oz option, we found no significant differences (p-value = `r result$p.value`).

- H<sub>o</sub> : No difference in ratings between the 4.5oz group and the 3oz group
- H<sub>a</sub> : Ratings in the 4.5oz group is higher than the 3oz group

```{r binom-printout} 
result
```

Because the ratings were not significantly higher in the 4.5oz option, we took the ratings from the 3oz option.

```{r filter-ranks}
ranks <- ranks %>% 
  filter(volume == 3) 
```

```{r join}
dat <- left_join(inks, ranks, by = c("Ink" = "name"))
```

## Noodler has an average of 4.5 out of 5 stars

In total, there are `r ranks %>% filter(volume == 3) %>% nrow()` unique inks listed in the Goulet Pen website; and on average, people gave Noodler's ink `r sprintf("%.1f", mean(ranks$rating))` out of 5 stars.

```{r plot-ratings}
dat %>%
  filter(!is.na(rating)) %>% 
  ggplot(aes(x = rating)) + 
    geom_bar() +
    ggtitle(sprintf("Noodler's (3oz) Ink Ratings from Goulet Pens (n=%s)",
                    nrow(ranks)))
```

The inks with 4.9+ stars are:

```{r table-top-inks}
dat %>% 
  filter(rating >= 4.9) %>% 
  arrange(desc(rating)) %>% 
  mutate(
    Ink = sprintf("[%s](http://www.gouletpens.com/search?query=%s)", Ink, itemId)
  ) %>% 
  head(20) %>% 
  select(Ink, Rating=rating, 
         Bulletproof, Eternal, Water_Resistant
         ) %>% 
  knitr::kable()
```

Unfortunately, none of the top rated inks are bulletproof, eternal, and/or water resistant.

## The top shortlisted inks

Here, I merged the Goulet Pen ratings with the shortlist of 43 inks and filtered for inks rated above average.

```{r table-my-top-inks}
dat %>% 
  filter(
    rating >= 4.5,
    Ink %in% inks_subset$Ink
    ) %>% 
  arrange(desc(rating)) %>% 
  mutate(
    rank = rank(desc(rating), ties.method="min"),
    Ink = sprintf("[%s](http://www.gouletpens.com/search?query=%s)", Ink, itemId)
  ) %>% 
  select(Rank = rank, 
         Ink,
         Rating = rating, 
         Bulletproof, Eternal, Water_Resistant) %>% 
  knitr::kable()
```

From this list, it looks like I should get "Black", "Blue Ghost", or "General of the Armies": Black is an unexceptional colour and is easily found in regular ballpoint pens; Blue Ghost is an invisible ink and I want something I can see; and General of the Armies is a maybe.

p.s.

I have already have "Air Corp Blue Black".
