---
title: "STAT545 enrollment"
author: "csiu"
date: "September 25, 2015"
output:
    html_document:
        keep_md: yes
---

As part of my education in graduate school, I'm enrolled in [STAT 545A: Exploratory Data Analysis](https://stat545-ubc.github.io). Here we learn data wrangling, exploration, and analysis with R.

```{r, echo=FALSE}
library(tidyr)
library(ggplot2)
library(grid)

# setwd('_drafts/2015-09-25-stat545-enrollment/')
data <- read.delim('data/2015_STAT545_enrollment.csv', sep=",")
```

The instructor, [\@JennyBryan](https://twitter.com/JennyBryan), recently compiled a CSV of [enrollment in STAT 545A in 2015 by degree and program](https://gist.github.com/jennybc/924fe242a31e0239762f#file-2015_stat545_enrollment-csv). And I decided to do some exploring of this data. Below is what the data looks like:

```{r, echo=FALSE, dev="png", dpi=340, out.width="604px"}
databysubject <- gather(data, Program, Count, -Subject)
ggplot(subset(databysubject, subset = (databysubject$Program != "Total")),
       aes(y=Subject, x=Program, fill=Count, label=Count)) +
  geom_tile(color="grey70") +
  geom_text(color="white") +
  xlab("") +
  ylab("") +
  scale_fill_gradient(low="white", high="blue") +
  #scale_y_discrete(limits = rev(levels(data$Subject))) +
  scale_y_discrete(limits = rev(as.character(data$Subject))) +
  guides(fill=FALSE) +
  theme(panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        axis.ticks.margin = unit(c(0.1,-1), 'lines'),
        axis.ticks = element_line(colour="grey90"),
        axis.text = element_text(colour="black"))
```

This heatmap is orded by number of students by program. Here we see that most students taking this course belong to Statistics.

