---
output:
    html_document:
        keep_md: yes
---
---
layout: post
author: "csiu"
date: 2016-05-06
title: "Posters submitted to ISMB"
categories: update
tags: data-analysis
---
```{r include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      results='hide',
                      fig.retina = 2,
                      fig.path='/Users/csiu/project/blog/myblog/img/figure/2016-05-06/')
options(knitr.table.format = 'markdown')
```

```{r load-packages, message=FALSE, warning=FALSE}
## Scraping HTML
library(rvest)

## Reformatting data
library(tidyr)
library(dplyr)
library(readr)

## Visualization
library(ggplot2)
library(viridis)
library(cowplot)

theme_set(theme_bw())
```

Yesterday I submitted a poster abstract to the [ISMB 2016 conference in Orlando, Florida](https://www.iscb.org/ismb2016). This submission -- submission #449 to category "Epigenetics"" -- was a late submission, which meant that the results of submissions from rounds 1 are done and accepted abstracts have been [released/listed](https://www.iscb.org/cms_addon/conferences/ismb2016/posterlist.php) onto the conference’s website.

*I was curious.*

My submission number is 449, this meant there were 448 submissions before mine. I want to know the following:

1. How many abstracts have been accepted during round 1?
2. Of the accepted abstracts in round 1, what proportion of them were categorized as "Epigenetic"?

## 16 possible categories

According to the [ISMB website](https://www.iscb.org/cms_addon/conferences/ismb2016/posterlist.php), there are 16 possible categories to which your abstract may belong.

```{r scrape-web, results='markup'}
## url
main_url <- "https://www.iscb.org/cms_addon/conferences/ismb2016/posterlist.php"


## getting the categories
categories <- read_html(main_url) %>% 
  html_nodes(".style1 .noBullet") %>% 
  html_text() 

(categories <- data.frame(categories = categories) %>%
  separate(col = categories, into = c("id", "category"), sep="\\)") %>%
  mutate(cat_url = sprintf("%s?cat=%s", main_url, id))) %>%
  mutate(
    Item = sprintf("%s.", row_number()),
    `Category/Link to abstracts` = sprintf("[%s](%s)", category, cat_url)
  ) %>%
  select(Item, `Category/Link to abstracts`, ID=id) %>%
  knitr::kable()


## helper to get the titles per category
get_titles <- function(cat_url){
  cat_url %>% #sprintf("%s?cat=%s", main_url, "D") %>%
  read_html() %>%
  html_nodes(".item-page .well") %>%
  html_text()
}
```

And if we assume that people submit equally to all categories, we expect each category to contain 6.25% (`100%/16`) of the total number of abstracts.

## Majority of posters are classified as "Sequence Analysis"

When we look at the number of abstracts to each category, we find the majority of them belong to "Sequence Analysis".

```{r calculations}
## number of posters per category
categories <- categories %>% 
  group_by(category) %>%
  mutate(
    poster = get_titles(cat_url) %>% list(),
    n_poster = poster %>% unlist() %>% length()
    ) %>% 
  ungroup()

## total number of posters
total_posters <- sum(categories$n_poster)
```

```{r poster-proportions, fig.height=5, fig.width=7.5}
categories %>%
  mutate(percent = sprintf("%.1f%%", n_poster/total_posters*100)) %>%
  ggplot(aes(x = reorder(category, n_poster),
             y = n_poster,
             label = percent)) +
    geom_bar(stat = "identity") +
    geom_text(y=99, hjust=1) +
    ylim(0,95) +
    xlab("") +
    ylab("Count") +
  coord_flip() +
    ggtitle(sprintf("Number of abstracts per category [n = %s]", total_posters)) +
    theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))
```

Here we also see -- as of Fri, May 6, 2016 -- there is a total of 366 abstracts accepted. From this total, only ~2.7% or 10 abstracts have been listed as "Epigenetics". 

*Not bad news.*

```{r write-data, eval=FALSE}
## The list of 366 abstracts
abstracts <- categories$poster
names(abstracts) <- categories$id

## to data frame
abstracts <- abstracts %>% 
  unlist()  %>% 
  as.data.frame() %>% 
  tbl_df()

colnames(abstracts) <- "title"

## make rowname into a column
abstracts$id <- rownames(abstracts) %>% 
  {sub("\\d+$", "", .)}
rownames(abstracts) <- NULL

abstracts %>% 
  select(category=id, title) %>%
  write_tsv("ismb-round1-accepted-abstracts.tsv")
```

p.s.

I have also made available the list of [366 abstracts](https://gist.github.com/csiu/b61c628478cfb57e2e7da91be7f88446) and the category to which they belong.